---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# easybio

<!-- badges: start -->
<!-- badges: end -->

You can install the development version of easybio from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("BiocManger")
BiocManger::install("snowGlint/easybio")
```

To know how to use this package, please see the [wiki](https://github.com/snowGlint/easybio/wiki).

## Coding Principles Reminders

- Use `[[` and `slot` to access child nodes instead of `$` or `@`.
- Ensure all operations are written in scripts, not executed directly in the terminal.
- Save all raw data used for your visualizations.
- Detailed script name to describle its target and number it by order.
- Use `lintr`.

An example project looks like this:

```r
r$> fs::dir_tree('example-project/')
example-project/
+-- .gitignore
+-- main.R
+-- patch1-GSEA4MF.R
+-- Rawdata
+-- README.qmd
\-- Result
    +-- main
    +-- patch1
    \-- README
```
